Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2025-04-18 full_version/leftlinear.tex
coerenza

	Then the following hold true:
	\begin{enumerate}
		\item if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and another derivation $\der{F}=\{\dder{F}_i\}_{i=0}^1$, then $\phi^{-1}_{G_0}*\der{E}*\phi_{G_2}$ is a pre-switch for $\der{F}$ along $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$;
		\item if $\{\psi_X\}_{X\in \Deltamin(\der{E})}$ is an abstraction equivalence between $\der{E}$ and $\der{G}=\{\dder{G}_{i}\}_{i=0}^1$, then $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$.
	\end{enumerate}
	\begin{remark}Recall that, by \Cref{rem:seqabs}, $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$ is an independence pair between $\dder{F}_0$ and $\dder{F}_1$.
	\end{remark}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item To prove this, it is enough to compute:
		\[\begin{split}
			m_{\der{F},0}=&\phi_{G_0}\circ m_0\\=&\phi_{G_0}\circ f'_0\circ i'_1
		\end{split} \quad \begin{split}\phi_{G_0}\circ	m'_0=&\phi_{G_0}\circ f_0\circ i_1\\=&f_{\der{F},0}\circ \phi_{D_{0}}\circ i_1
	 \end{split} 
	\]
	\[\begin{split}
	h_{\der{F},1}=&\phi_{G_2}\circ h_1\\=&\phi_{G_2}\circ g_1' \circ i_0'
	\end{split} \qquad \begin{split}
		\phi_{G_2}\circ h'_1=&\phi_{G_2}\circ g_1\circ i_0\\=&g_{\der{F}, 1}\circ \phi_{D_1}\circ i_0
	\end{split}\]

@@

\begin{corollary}\label{lem:cose}Let
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation and suppose
	that $(i_0, i_1)$ is an independence pair between $\dder{D}_0$
	and $\der{D}_1$.  If $\der{E}=\{\dder{E}_i\}_{i=0}^1$ is a
	pre-switch of $\der{D}$ along $(i_0, i_1)$, then there is an arrow
	$z_0\colon P\to D_{\der{E},0}$ in $\mathcal{M}$ fitting in the square below.
	\[\xymatrix{P \ar@{.>}[r]^-{z_0}  \ar@{>->}[d]_{p_0}& D_{\der{E}, 0} \ar@{>->}[d]^{f_{\der{E},0}} \\ D_0 \ar@{>->}[r]_{f_0} & G_0 }\]
\end{corollary}
\begin{proof}
	Consider the arrow $u_1\colon K_1\to P$ obtained using
	\Cref{prop:tec}. It fits in the big square on the left
	\[		\xymatrix{K_1 \ar[r]^{u_1} \ar@{>->}[d]_{l_1}&P
		\ar[r]^{\id{P}}\ar@{>->}[d]_{p_0}& P\ar@{>->}[d]^{p_0}\\L_1
		\ar[d]_{\id{L_1}}\ar[r]^{i_1}&D_0 \ar[r]^{\id{D_0}}
		\ar[d]_{\id{D_0}}& D_{0} \ar@{>->}[d]^{f_0}&K_1 \ar@{>->}[d]_{l_1}\ar[r]^{k_{\der{E},0}}&
		D_{\der{E},0} \ar@{>->}[d]^{f_{\der{E},0}} & K_1 \ar[r]^{u_1}
		\ar@{>->}[d]_{l_1}& P \ar@{>->}[d]^{f_0\circ
			p_0}\\
		L_1\ar[r]^{i_1} \ar@/_.4cm/[rr]_{m_{\der{E},0}}&D_0
		\ar@{>->}[r]^{f_0}&G_0 & L_1\ar[r]_{m_{\der{E},0}}&G_0 &L_1
		\ar[r]_{m_{\der{E},0}}&G_0}\]

	Since $f_0$ is a monomorphism, all the inner squares of the left diagram are pullbacks, thus the whole big square is a pullback too.
	Now the thesis follows applying \Cref{lem:radj} to the pair of squares on the right.
	\end{proof}

\begin{example}\label{ex:abs2}
Let $\der{D}$ be the derivation of \Cref{ex:abs} and consider the pre-switch $\der{F}$. Then we can build the diagram belo
	\[\xymatrix@C=20pt@R=20pt{&&1 \ar@{>->}@/_.9cm/[ddrr]_(.2){\delta^3_0}|(.72)\hole
	\ar@{>->}[dr]^{\delta^3_0}&& 1\ar@{>->}@/^.9cm/[ddll]^(.2){\delta^3_0}
	\ar@{>->}[dl]_{\delta^3_0}\\
	&1\ar@{>->}[dr]^{\delta^3_0}\ar@{>->}[dl]_{\id{1}}
	\ar@{>->}@/_.7cm/[ddrr]|(.38)\hole^(.7){\delta^3_0}\ar@{>->}[ur]^{\id{1}}&& 3 &&
	1
	\ar@{>->}@/^.7cm/[ddll]|(.38)\hole_(.7){\delta^3_0}\ar@{>->}[dl]_{\delta^3_0}\ar@{>->}[lu]_{\delta^3_0}
	\ar@{>->}[dr]^{\delta^2_0}\\
	1 \ar@/_.7cm/[ddrr]_(.2){\delta^3_0}|(.325)\hole|(.81)\hole
	\ar@{>->}[dr]^(.35){\delta^3_0}|(.6)\hole && 3
	\ar@{>->}[dl]|(.4)\hole^(.65){\id{3}}\ar@{>->}[ur]|(.5)\hole^(.7){\id{3} }&&3
	\ar@{>->}[dr]|(.4)\hole_(.65){i_3}
	\ar@{>->}[ul]|(.475)\hole_(.7){\id{3}}&&2\ar@{>->}@/^.7cm/[ddll]^(.2){i_2}|(.325)\hole|(.81)\hole\ar@{>->}[dl]_(.35){i_2}|(.6)\hole\\
	&3 &&3	\ar@{>->}[dl]^{\id{3}}\ar@{>->}[ur]^{\id{3}}\ar@{>->}[ul]_{\id{3}}&&4\\
	1	\ar@{>->}@/^.7cm/[uurr]^(.2){\delta^3_0} \ar@{>->}[ur]_(.35){\delta^3_0}|(.61)\hole && 3	\ar@{>->}[ul]|(.4)\hole_(.65){\id{3}}\ar@{>->}[dr]|(.5)\hole_(.65){i_3}	&&4\ar@{>->}[ur]|(.4)\hole^(.65){t} \ar@{>->}[dl]|(.5)\hole^(.65){\id{4}}	&& 1\ar[ul]^(.35){\delta^4_0}|(.61)\hole\ar@/_.7cm/[uull]_(.2){\delta^3_0}\\ 
	&1	\ar@{>->}[ur]_{\delta^3_0} \ar@{>->}[dr]_{i_2}	\ar@{>->}[ul]^{\id{1}}\ar@/^.7cm/[uurr]_(.65){\delta^3_0}&&4&& 1\ar@{>->}[ur]_{\id{1}} \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}[dl]^{\id{1}} \ar@{>->}@/_.7cm/[uull]^(.65){\delta^3_0}\\&& 2	\ar@{>->}[ur]_{i_2}\ar@{>->}@/^.9cm/[uurr]^(.2){i_2}&& 1 \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}@/_.9cm/[uull]_(.2){\delta^3_0} |(.72)\hole }\]

@@

\begin{definition}[Switch] 	Let $(\X, \R)$ be a left-linear rewriting system, with $\X$ an $\mathcal{M}$-adhesive category.
	Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram below, where the cone with vertex $P$ is a pullback and $u_0\colon K_0\to P$ and $u_1\colon K_1\to P$ are the arrows built in \Cref{prop:tec}.
	\[\xymatrix@R=15pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0 \ar @/^.35cm/[ddrr]_(.2){u_0}
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}|(.68)\hole
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}|(.68)\hole 
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar @/_.35cm/[ddll]^(.2){u_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}|(.7)\hole&& G_1 && \ar@{>->}[ll]^{f_1}|(.7)\hole D_1
		\ar[rr]_{g_1}&& G_2 \\ & & & & P \ar[urr]_{p_1} \ar@{>->}[ull]^{p_0}}
	\]

	Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$, depicted below, be a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=15pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	We say that $\der{E}$ is a \emph{switch} for $\der{D}$ along $(i_0, i_1)$ if there is an arrow $z_1\colon P\to D'_1$ fitting in the diagram below, where $z_0\colon P\to D'_0$ is the arrow constructed in \Cref{lem:cose}.
	\[\xymatrix@R=18pt{D'_0  \ar[d]_{g'_0}& P \ar[r]^{p_1} \ar[l]_{z_0} \ar@{.>}[d]^{z_1} & D_1 \ar[d]^{g_1}\\ G'_1 & D'_1 \ar[r]_{g'_1} \ar@{>->}[l]^{f'_1} & G_2}\]

@@

\begin{remark}\label{rem:quadrati}
Let $\der{D}$ be a derivation made by two sequentially independence direct derivations with independence pair $(i_0, i_1)$. Suppose that $\der{E}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$. Then we have the solid part of the diagram below. If, moreover $\der{E}$ is a switch, then we can complete it with the dotted arrow $z_1\colon P\to D'_1$. 
		\[\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){i_0}|(.7)\hole
		\ar[dr]^{h_0}&& L_1\ar@/^.8cm/[ddll]^(.2){i_1}
		\ar[dl]_{m_1}\\&K_0\ar[dr]^{k_0}\ar@{>->}[dl]_{l_0}
		\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){u_0}\ar[ur]^{r_0}&& G_1 &&
		K_1
		\ar@/^.6cm/[ddll]|(.36)\hole_(.65){u_1}\ar[dl]_{k_1}\ar@{>->}[lu]_{l_1}
		\ar[dr]^{r_1}\\L_0
		\ar@/_.6cm/[ddrr]_(.2){i'_1}|(.31)\hole|(.81)\hole
		\ar[dr]^(.4){m_0}|(.61)\hole && D_0
		\ar@{>->}[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&D_1
		\ar[dr]|(.4)\hole^(.65){g_1}
		\ar@{>->}[ul]|(.5)\hole_(.65){f_1}&&R_1\ar@/^.6cm/[ddll]^(.2){i'_0}|(.31)\hole|(.81)\hole\ar[dl]_{h_1}|(.6)\hole\\&G_0
		&&P\ar@{.>}[dr]_{z_1}	\ar@{>}[dl]^{z_0}\ar[ur]^(.4){p_1}\ar@{>->}[ul]_(.4){p_0}&&G_2\\L_1	\ar@/^.6cm/[uurr]^(.2){i_1} \ar[ur]_(.35){m'_0}|(.61)\hole && D'_0	\ar@{>->}[ul]|(.4)\hole_(.65){f'_0}\ar[dr]|(.5)\hole_(.65){g'_0}	&&D'_1\ar[ur]|(.4)\hole^(.65){g'_1} \ar@{>->}[dl]|(.5)\hole^(.65){f'_1}	&& R_0 \ar[ul]^(.35){h'_1}|(.61)\hole\ar@/_.6cm/[uull]_(.2){i_0}\\ &K_1	\ar[ur]_{k'_0} \ar[dr]_{r_1}	\ar@{>->}[ul]^{l_1}\ar@/^.6cm/[uurr]_(.65){u_1}&&G'_1&& K_0\ar[ur]_{r_0} \ar[ul]^{k'_1} \ar@{>->}[dl]^{l_0} \ar@/_.6cm/[uull]^(.65){u_0}\\&& R_1	\ar[ur]_{h'_0}\ar@/^.8cm/[uurr]^(.2){i'_0}&& L_0 \ar[ul]^{m'_1} \ar@/_.8cm/[uull]_(.2){i'_1} |(.69)\hole }\] 
\end{remark}

\begin{remark}\label{rem:un}
	Notice that, since $f'_1$ belongs to $\mathcal{M}$, there is at most one arrow $z_1\colon P\to D'_1$ such that $f'{1}\circ z_1=g'_0\circ z_0$.
\end{remark}

@@

	\smallskip \noindent $(2\Rightarrow 1)$. By hypothesis the solid part of the diagram below commutes.
	\[\xymatrix@C=30pt{P \ar[r]^{z_0} \ar@{.>}[dr]^{z_1}\ar[d]_{p_1} & D'_0 \ar[r]^{g'_0} & G'_1 \ar@{>->}[dr]^{\iota'_{1, G'_1}}\\ D_1 \ar[dr]_{g'_1}& D'_1 \ar[d]^{g'_1}\ar@{>->}[r]^{f'_1} &G'_1 \ar@{>->}[r]^{\iota'_{1, G'_1}} &\lpro \dder{E}_0 \rpro 	\ar@{>->}[d]^{p_1} \\ & G'_1 	\ar[rr]_{\iota_{\der{E}, G_2}}&& \tpro{E} }\]

	By \Cref{rem:zero1} the inner square is a pullback, proving that the dotted arrow $z_1\colon P\to D'_1$ exists.  Since, by \Cref{lem:colim} $\iota'_{1, G'_1}$ is in $\mathcal{M}$, we conclude that $f'\circ z_1=g'\circ z_0$ as wanted.
\end{proof}
\iffalse 
We turn now to simmetry of the switchability relation. Surely one desires an analog of \Cref{rem:switch} even for switches: this is assured by the following proposition.

\begin{proposition}\label{prop:pb}
	Let $\der{D}:=\{\dder{D}_{i}\}_{i=0}^1$ be the derivation below, with $(i_0, i_1)$ an independence pair between $\dder{D}_0$  and $\dder{D}_1$.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]

	If $\der{E}:=\{\dder{E}_{i}\}_{i=0}^1$ is a switch for $\dder{D}$ along $(i_0, i_1)$, then the square below is a pullback.
	\[\xymatrix{P  \ar[r]^-{z_1} \ar[d]_{z_0}& D_{\der{E}, 1} \ar@{>->}[d]^{f_{\der{E},1}} \\ D_{\der{E},0} \ar[r]_{g_{\der{E},0}}& G_{\der{E},1}}\]

>>>>

	Then the following hold true:
	\begin{enumerate}
		\item if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and $\der{F}=\{\dder{F}_i\}_{i=0}^1$, then $\phi^{-1}_{G_0}*\der{E}*\phi_{G_2}$ is a pre-switch for $\der{F}$ along $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$;
		\item if $\{\psi_X\}_{X\in \Deltamin(\der{E})}$ is an abstraction equivalence between $\der{E}$ and $\der{G}=\{\dder{G}_{i}\}_{i=0}^1$, then $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$.
	\end{enumerate}
	\begin{remark}Recall that, by \Cref{rem:seqabs}, $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$ is an independence pair between $\dder{F}_0$ and $\dder{F}_1$.
	\end{remark}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item To prove this, it is enough to compute:
		\[\begin{split}
			m_{\der{F},0}=&\phi_{G_0}\circ m_0\\=&\phi_{G_0}\circ f'_0\circ i'_1
		\end{split} \quad \begin{split}\phi_{G_0}\circ	m'_0=&\phi_{G_0}\circ f_0\circ i_1\\=&f_{\der{F},0}\circ \phi_{D_{0}}\circ i_1
	 \end{split} 
	\]
	\[\begin{split}
	h_{\der{F},1}=&\phi_{G_2}\circ h_1\\=&\phi_{G_2}\circ g_1' \circ i_0'
	\end{split} \qquad \begin{split}
		\phi_{G_2}\circ h'_1=&\phi_{G_2}\circ g_1\circ i_0\\=&g_{\der{F}, 1}\circ \phi_{D_1}\circ i_0
	\end{split}\]

@@

\begin{corollary}\label{lem:cose}Let
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation and suppose
	that $(i_0, i_1)$ is an independence pair between $\dder{D}_0$
	and $\der{D}_1$.  If $\der{E}=\{\dder{E}_i\}_{i=0}^1$ is a
	pre-switch of $\der{D}$ along $(i_0, i_1)$, then there is a unique arrow
	$z_0\colon P\to D_{\der{E},0}$ fitting in the square below. Moreover, such arrow is in $\mathcal{M}$.
	\[\xymatrix{P \ar@{>.>}[r]^-{z_0}  \ar@{>->}[d]_{p_0}& D_{\der{E}, 0} \ar@{>->}[d]^{f_{\der{E},0}} \\ D_0 \ar@{>->}[r]_{f_0} & G_0 }\]
\end{corollary}
\begin{proof}
	Consider the arrow $u_1\colon K_1\to P$ obtained using
	\Cref{prop:tec}. It fits in the big square on the left
	\[		\xymatrix{K_1 \ar[r]^{u_1} \ar@{>->}[d]_{l_1}&P
		\ar[r]^{\id{P}}\ar@{>->}[d]_{p_0}& P\ar@{>->}[d]^{p_0}\\L_1
		\ar[d]_{\id{L_1}}\ar[r]^{i_1}&D_0 \ar[r]^{\id{D_0}}
		\ar[d]_{\id{D_0}}& D_{0} \ar@{>->}[d]^{f_0}&K_1 \ar@{>->}[d]_{l_1}\ar[r]^{k_{\der{E},0}}&
		D_{\der{E},0} \ar@{>->}[d]^{f_{\der{E},0}} & K_1 \ar[r]^{u_1}
		\ar@{>->}[d]_{l_1}& P \ar@{>->}[d]^{f_0\circ
			p_0}\\
		L_1\ar[r]^{i_1} \ar@/_.4cm/[rr]_{m_{\der{E},0}}&D_0
		\ar@{>->}[r]^{f_0}&G_0 & L_1\ar[r]_{m_{\der{E},0}}&G_0 &L_1
		\ar[r]_{m_{\der{E},0}}&G_0}\]

	Since $f_0$ is a monomorphism, all the inner squares of the left diagram are pullbacks, thus the whole big square is a pullback too.
	Now the thesis follows applying \Cref{lem:radj} to the pair of squares on the right. Since $f_0$ and $p_0$ are element of $\mathcal{M}$, then $f_{\der{E},0}\circ z_0$ is in $\mathcal{M}$ too and so $z_0$ is in $\mathcal{M}$ by \Cref{prop:deco}.	
	\end{proof}

\begin{example}\label{ex:abs2}
Let $\der{D}$ be the derivation of \Cref{ex:abs} and consider the pre-switch $\der{F}$. Then we can build the diagram below
	\[\xymatrix@C=20pt@R=20pt{&&1 \ar@{>->}@/_.9cm/[ddrr]_(.2){\delta^3_0}|(.72)\hole
	\ar@{>->}[dr]^{\delta^3_0}&& 1\ar@{>->}@/^.9cm/[ddll]^(.2){\delta^3_0}
	\ar@{>->}[dl]_{\delta^3_0}\\
	&1\ar@{>->}[dr]^{\delta^3_0}\ar@{>->}[dl]_{\id{1}}
	\ar@{>->}@/_.7cm/[ddrr]|(.38)\hole^(.7){\delta^3_0}\ar@{>->}[ur]^{\id{1}}&& 3 &&
	1
	\ar@{>->}@/^.7cm/[ddll]|(.38)\hole_(.7){\delta^3_0}\ar@{>->}[dl]_{\delta^3_0}\ar@{>->}[lu]_{\delta^3_0}
	\ar@{>->}[dr]^{\delta^2_0}\\
	1 \ar@/_.7cm/[ddrr]_(.2){\delta^3_0}|(.325)\hole|(.81)\hole
	\ar@{>->}[dr]^(.35){\delta^3_0}|(.6)\hole && 3
	\ar@{>->}[dl]|(.4)\hole^(.65){\id{3}}\ar@{>->}[ur]|(.5)\hole^(.7){\id{3} }&&3
	\ar@{>->}[dr]|(.4)\hole_(.65){i_3}
	\ar@{>->}[ul]|(.475)\hole_(.7){\id{3}}&&2\ar@{>->}@/^.7cm/[ddll]^(.2){i_2}|(.325)\hole|(.81)\hole\ar@{>->}[dl]_(.35){i_2}|(.6)\hole\\
	&3 &&3	\ar@{>->}[dl]^{\id{3}}\ar@{>->}[ur]^{\id{3}}\ar@{>->}[ul]_{\id{3}}&&4\\
	1	\ar@{>->}@/^.7cm/[uurr]^(.2){\delta^3_0} \ar@{>->}[ur]_(.35){\delta^3_0}|(.61)\hole && 3	\ar@{>->}[ul]|(.4)\hole_(.65){\id{3}}\ar@{>->}[dr]|(.5)\hole_(.65){i_3}	&&4\ar@{>->}[ur]|(.4)\hole^(.65){t} \ar@{>->}[dl]|(.5)\hole^(.65){\id{4}}	&& 1\ar[ul]^(.35){\delta^4_0}|(.61)\hole\ar@/_.7cm/[uull]_(.2){\delta^3_0}\\ 
	&1	\ar@{>->}[ur]_{\delta^3_0} \ar@{>->}[dr]_{i_2}	\ar@{>->}[ul]^{\id{1}}\ar@/^.7cm/[uurr]_(.65){\delta^3_0}&&4&& 1\ar@{>->}[ur]_{\id{1}} \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}[dl]^{\id{1}} \ar@{>->}@/_.7cm/[uull]^(.65){\delta^3_0}\\&& 2	\ar@{>->}[ur]_{i_2}\ar@{>->}@/^.9cm/[uurr]^(.2){i_2}&& 1 \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}@/_.9cm/[uull]_(.2){\delta^3_0} |(.72)\hole }\]

@@

\begin{definition}[Switch] 	Let $(\X, \R)$ be a left-linear rewriting system, with $\X$ an $\mathcal{M}$-adhesive category.
	Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram below, where the cone with vertex $P$ is a pullback and $u_0\colon K_0\to P$ and $u_1\colon K_1\to P$ are the arrows built in \Cref{prop:tec}.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0 \ar @/^.35cm/[ddrr]_(.2){u_0}
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}|(.68)\hole
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}|(.68)\hole 
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar @/_.35cm/[ddll]^(.2){u_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}|(.7)\hole&& G_1 && \ar@{>->}[ll]^{f_1}|(.7)\hole D_1
		\ar[rr]_{g_1}&& G_2 \\ & & & & P \ar[urr]_{p_1} \ar@{>->}[ull]^{p_0}}
	\]

	Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$, depicted below, be a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	We say that $\der{E}$ is a \emph{switch} for $\der{D}$ along $(i_0, i_1)$ if there is an arrow $z_1\colon P\to D'_1$ fitting in the diagram below, where $z_0\colon P\to D'_0$ is the arrow constructed in \Cref{lem:cose}.
	\[\xymatrix@R=18pt{D'_0  \ar[d]_{g'_0}& P \ar[r]^{p_1} \ar@{>->}[l]_{z_0} \ar@{.>}[d]^{z_1} & D_1 \ar[d]^{g_1}\\ G'_1 & D'_1 \ar[r]_{g'_1} \ar@{>->}[l]^{f'_1} & G_2}\]

@@

\begin{remark}\label{rem:quadrati}
Let $\der{D}$ be a derivation made by two sequentially independence direct derivations with independence pair $(i_0, i_1)$. Suppose that $\der{E}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$. Then we have the solid part of the diagram below. If, moreover $\der{E}$ is a switch, then we can complete it with the dotted arrow $z_1\colon P\to D'_1$. 
		\[\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){i_0}|(.7)\hole
		\ar[dr]^{h_0}&& L_1\ar@/^.8cm/[ddll]^(.2){i_1}
		\ar[dl]_{m_1}\\&K_0\ar[dr]^{k_0}\ar@{>->}[dl]_{l_0}
		\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){u_0}\ar[ur]^{r_0}&& G_1 &&
		K_1
		\ar@/^.6cm/[ddll]|(.36)\hole_(.65){u_1}\ar[dl]_{k_1}\ar@{>->}[lu]_{l_1}
		\ar[dr]^{r_1}\\L_0
		\ar@/_.6cm/[ddrr]_(.2){i'_1}|(.31)\hole|(.81)\hole
		\ar[dr]^(.4){m_0}|(.61)\hole && D_0
		\ar@{>->}[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&D_1
		\ar[dr]|(.4)\hole^(.65){g_1}
		\ar@{>->}[ul]|(.5)\hole_(.65){f_1}&&R_1\ar@/^.6cm/[ddll]^(.2){i'_0}|(.31)\hole|(.81)\hole\ar[dl]_{h_1}|(.6)\hole\\&G_0
		&&P\ar@{.>}[dr]_{z_1}	\ar@{>->}[dl]^{z_0}\ar[ur]^(.4){p_1}\ar@{>->}[ul]_(.4){p_0}&&G_2\\L_1	\ar@/^.6cm/[uurr]^(.2){i_1} \ar[ur]_(.35){m'_0}|(.61)\hole && D'_0	\ar@{>->}[ul]|(.4)\hole_(.65){f'_0}\ar[dr]|(.5)\hole_(.65){g'_0}	&&D'_1\ar[ur]|(.4)\hole^(.65){g'_1} \ar@{>->}[dl]|(.5)\hole^(.65){f'_1}	&& R_0 \ar[ul]^(.35){h'_1}|(.61)\hole\ar@/_.6cm/[uull]_(.2){i_0}\\ &K_1	\ar[ur]_{k'_0} \ar[dr]_{r_1}	\ar@{>->}[ul]^{l_1}\ar@/^.6cm/[uurr]_(.65){u_1}&&G'_1&& K_0\ar[ur]_{r_0} \ar[ul]^{k'_1} \ar@{>->}[dl]^{l_0} \ar@/_.6cm/[uull]^(.65){u_0}\\&& R_1	\ar[ur]_{h'_0}\ar@/^.8cm/[uurr]^(.2){i'_0}&& L_0 \ar[ul]^{m'_1} \ar@/_.8cm/[uull]_(.2){i'_1} |(.69)\hole }\] 
\end{remark}

\begin{remark}\label{rem:un}
	Notice that, since $f'_1$ belongs to $\mathcal{M}$, there is at most one arrow $z_1\colon P\to D'_1$ such that $f'_{1}\circ z_1=g'_0\circ z_0$.
\end{remark}

@@

	\smallskip \noindent $(2\Rightarrow 1)$. By hypothesis the solid part of the diagram below commutes.
	\[\xymatrix@C=30pt{P \ar@{>->}[r]^{z_0} \ar@{.>}[dr]^{z_1}\ar[d]_{p_1} & D'_0 \ar[r]^{g'_0} & G'_1 \ar@{>->}[dr]^{\iota'_{1, G'_1}}\\ D_1 \ar[dr]_{g'_1}& D'_1 \ar[d]^{g'_1}\ar@{>->}[r]^{f'_1} &G'_1 \ar@{>->}[r]^{\iota'_{1, G'_1}} &\lpro \dder{E}_0 \rpro 	\ar@{>->}[d]^{p_1} \\ & G'_1 	\ar[rr]_{\iota_{\der{E}, G_2}}&& \tpro{E} }\]

	By \Cref{rem:zero1} the inner square is a pullback, proving that the dotted arrow $z_1\colon P\to D'_1$ exists.  Since, by \Cref{lem:colim} $\iota'_{1, G'_1}$ is in $\mathcal{M}$, we conclude that $f'\circ z_1=g'\circ z_0$ as wanted.
\end{proof}

By \Cref{thm:switch_uni} and \Cref{ex:abs2} we know that being a switch is not a property invariant under  abstraction equivalence. Nonetheless we  can still prove an analog to \Cref{lem:abs}.
